---
title: Calculate Correlation
summary:  This notebook is to calculate and display statisitcal significance.
output: html_notebook
---


```{r}
library(tidyverse)
library(dplyr)
library(ggpubr)
```

Import data from CSV.  See /wrangle/survey_Wrangle.Rmd for code used to 
create the CSV.

```{r}
HAB_MRIP <- read.csv("../data_out/sarasota.csv")
HAB_MRIP_small <- read.csv("../data_out/sarasota_year_month.csv")
```




```{r}
summary(HAB_MRIP_small$mean_brevis_count)
summary(HAB_MRIP_small$catch_count)


```

```{r}
#https://www.statmethods.net/graphs/density.html
# Add a Normal Curve (Thanks to Peter Dalgaard)
x <- HAB_MRIP_small$catch_count
h<-hist(x, breaks=10, col="red", xlab="MRIP Catch Count",
   main="MRIP Catch Distribution")
xfit<-seq(min(x),max(x),length=40)
yfit<-dnorm(xfit,mean=mean(x),sd=sd(x))
yfit <- yfit*diff(h$mids[1:2])*length(x)
lines(xfit, yfit, col="blue", lwd=2) 


```
```{r}
x <- HAB_MRIP_small$brevis_count
h<-hist(x, breaks=10, col="red", xlab="HAB Mean Brevis Count",
   main="HAB Distribution")
xfit<-seq(min(x),max(x),length=40)
yfit<-dnorm(xfit,mean=mean(x),sd=sd(x))
yfit <- yfit*diff(h$mids[1:2])*length(x)
lines(xfit, yfit, col="blue", lwd=2) 
```
```{r}
count(HAB_MRIP_small %>% filter(brevis_count<1000))
```


```{r}
head(HAB_MRIP_small)

#Plot Correltation coeffecient

#http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r

ggscatter(HAB_MRIP_small, x = "brevis_count", y = "catch_count", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "K. brevis", ylab = "Game Fish Caught")

#http://www.sthda.com/english/wiki/wiki.php?id_contents=7312

# x and y are numeric vectors with the same length


remove_outliers <- HAB_MRIP_small %>% filter(catch_count < 3000)

#Test 1
cor.test(remove_outliers$brevis_count, remove_outliers$catch_count, method=c("pearson", "kendall", "spearman"))
```


```{r}
#test 2.  mean_brevis_count
cor.test(remove_outliers$mean_brevis_count, remove_outliers$catch_count, method=c("pearson", "kendall", "spearman"))  #nearly the same
```


```{r}
#Test previous month 
cor.test(remove_outliers$previous_month_mean_brevis, remove_outliers$catch_count, method=c("pearson", "kendall", "spearman"))  #nearly the same

```
```{r}
#Test without high catch count
tight <- HAB_MRIP_small %>% filter(catch_count < 600,brevis_count > 0)
cor.test(tight$mean_brevis, tight$catch_count, method=c("pearson", "kendall", "spearman"))
```

Looking at the calculations above, we see virtually no difference between total 
brevis and mean brevis.  This means the number of monitors doesn't make 
much difference at least not over the course of a month.  

Using previous month brevis, the correlation gets worse meaning there is no
delayed reaction to a red tide event.  At least not month over month.  
If we had daily or weekly catch count data, we might see a better correlation 
week over week, but we don't have this data.  


```{r}
#Plot Correlation coeffecient

#http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r

ggscatter(remove_outliers, x = "brevis_count", y = "catch_count", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Aggregate K. brevis", ylab = "Game Fish Caught", title="Pearson's Correllation - Catch vs Brevis",
          add.params = list(color = "blue", fill = "lightgray"),cor.coef.coord =c(102821633,2000))

```
There is obviously a wide distribution of fish catch when K. brevis is zero.
This signifies other factors which impact fish catch more than red tide.  
However, I also notice that as K. brevis increases, there are few data points 
with high fish catch counts.  

What if we eliminate data where brevis count = 0? 

```{r}

summary(HAB_MRIP_small$catch_count)
summary(HAB_MRIP_small$brevis_count)

#ggplot(HAB_MRIP_small, aes(x=catch_count)) + geom_histogram(binwidth=.5)

remove_outliers <- HAB_MRIP_small %>% filter(catch_count < 600,brevis_count > 0)
head(remove_outliers)


#http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r

#Plot Correlation coeffecient

#http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r

ggscatter(remove_outliers, x = "mean_brevis_count", y = "catch_count", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Mean K. brevis", ylab = "Game Fish Caught", title="Pearson's Correllation - Catch vs Brevis",
          add.params = list(color = "blue", fill = "lightgray"),cor.coef.coord =c(4999000,500))
```

```{r}
head(HAB_MRIP_small)
medium_to_high_brevis <- HAB_MRIP_small %>% filter(mean_brevis_count > 100000)
head(medium_to_high_brevis)

#http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r

ggscatter(medium_to_high_brevis, x = "brevis_count", y = "catch_count", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "K. brevis", ylab = "Game Fish Caught")

```

