---
title: Calculate Correlation
summary:  This notebook is to calculate and display statisitcal significance.
output: html_notebook
---


```{r}
library(tidyverse)
library(dplyr)
library(ggpubr)
```

Import data from CSV.  See /wrangle/survey_Wrangle.Rmd for code used to 
create the CSV.

```{r}
HAB_MRIP <- read.csv("../data_out/sarasota.csv")
HAB_MRIP_small <- read.csv("../data_out/sarasota_year_month.csv")
```

```{r}
head(HAB_MRIP_small)

#http://www.sthda.com/english/wiki/wiki.php?id_contents=7312

# x and y are numeric vectors with the same length
cor.test(HAB_MRIP_small$brevis_count, HAB_MRIP_small$catch_count, method=c("pearson", "kendall", "spearman"))

remove_outliers <- HAB_MRIP_small %>% filter(catch_count < 600)

cor.test(remove_outliers$brevis_count, remove_outliers$catch_count, method=c("pearson", "kendall", "spearman"))

#test mean_brevis_count
cor.test(HAB_MRIP_small$mean_brevis_count, HAB_MRIP_small$catch_count, method=c("pearson", "kendall", "spearman"))  #nearly the same

#Test previous month total and mean
cor.test(HAB_MRIP_small$previous_month_brevis, HAB_MRIP_small$catch_count, method=c("pearson", "kendall", "spearman"))  #nearly the same
cor.test(HAB_MRIP_small$previous_month_mean_brevis, HAB_MRIP_small$catch_count, method=c("pearson", "kendall", "spearman"))  #nearly the same


```

Looking at the calculations above, we see virtually no difference between total 
brevis and mean brevis.  This means the number of monitors doesn't make 
much difference at least not over the course of a month.  

Using previous month brevis, the correlation gets worse meaning there is no
delayed reaction to a red tide event.  At least not month over month.  
If we had daily or weekly catch count data, we might see a better correlation 
week over week, but we don't have this data.  


```{r}
#Plot Correltation coeffecient

#http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r

ggscatter(HAB_MRIP_small, x = "brevis_count", y = "catch_count", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "K. brevis", ylab = "Game Fish Caught")

```
There is obviously a wide distribution of fish catch when K. brevis is zero.
This signifies other factors which impact fish catch more than red tide.  
However, I also notice that as K. brevis increases, there are few data points 
with high fish catch counts.  

What if we eliminate data where brevis count = 0? 

```{r}

summary(HAB_MRIP_small$catch_count)
summary(HAB_MRIP_small$brevis_count)

#ggplot(HAB_MRIP_small, aes(x=catch_count)) + geom_histogram(binwidth=.5)

remove_outliers <- HAB_MRIP_small %>% filter(catch_count < 600)



#http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r

ggscatter(remove_outliers, x = "brevis_count", y = "catch_count", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson", cor.coef.coord =c(400000000,400),
          xlab = "K. brevis", ylab = "Game Fish Caught")

```

```{r}
head(HAB_MRIP_small)
medium_to_high_brevis <- HAB_MRIP_small %>% filter(mean_brevis_count > 100000)
head(medium_to_high_brevis)

#http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r

ggscatter(medium_to_high_brevis, x = "brevis_count", y = "catch_count", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "K. brevis", ylab = "Game Fish Caught")

```

