---
title: Calculate Correlation
summary:  This notebook is to calculate and display statisitcal significance.
output: html_notebook
---


```{r}
library(tidyverse)
library(dplyr)
library(ggpubr)
```

Import data from CSV.  See /wrangle/survey_Wrangle.Rmd for code used to 
create the CSV.

```{r}
HAB_MRIP <- read.csv("../data_out/sarasota.csv")
HAB_MRIP_small <- read.csv("../data_out/sarasota_year_month.csv")
```




```{r}
summary(HAB_MRIP_small$mean_brevis_count)
summary(HAB_MRIP_small$catch_count)


```

```{r}
#https://www.statmethods.net/graphs/density.html
# Add a Normal Curve (Thanks to Peter Dalgaard)
x <- HAB_MRIP_small$catch_count
h<-hist(x, breaks=10, col="red", xlab="MRIP Catch Count",
   main="MRIP Catch Distribution")
xfit<-seq(min(x),max(x),length=40)
yfit<-dnorm(xfit,mean=mean(x),sd=sd(x))
yfit <- yfit*diff(h$mids[1:2])*length(x)
lines(xfit, yfit, col="blue", lwd=2) 


```
```{r}
x <- HAB_MRIP_small$mean_brevis_count
h<-hist(x, breaks=10, col="red", xlab="HAB Mean Brevis Count",
   main="HAB Distribution")
xfit<-seq(min(x),max(x),length=40)
yfit<-dnorm(xfit,mean=mean(x),sd=sd(x))
yfit <- yfit*diff(h$mids[1:2])*length(x)
lines(xfit, yfit, col="blue", lwd=2) 
```
```{r}
count(HAB_MRIP_small %>% filter(mean_brevis_count<1000))
count(HAB_MRIP_small %>% filter(mean_brevis_count>100000))
count(HAB_MRIP_small)
```


```{r}
remove_outliers <- HAB_MRIP_small %>% filter(catch_count < 600)

#Test 1
cor.test(remove_outliers$mean_brevis_count, remove_outliers$catch_count, method=c("pearson", "kendall", "spearman"))
```


```{r}
#Test previous month 
cor.test(remove_outliers$previous_month_mean_brevis, remove_outliers$catch_count, method=c("pearson", "kendall", "spearman"))  #nearly the same

```

```{r}
#Plot Correlation coeffecient

#http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r

ggscatter(remove_outliers, x = "mean_brevis_count", y = "catch_count", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Mean K. brevis", ylab = "Game Fish Caught", title="Pearson's Correllation - Catch vs Brevis",
          add.params = list(color = "blue", fill = "lightgray"),cor.coef.coord =c(3999999,500))

```
There is obviously a wide distribution of fish catch when K. brevis is zero.
This signifies other factors which impact fish catch more than red tide.  
However, I also notice that as K. brevis increases, there are few data points 
with high fish catch counts.  

What if we eliminate data where brevis count = 0? 

```{r}

summary(HAB_MRIP_small$catch_count)
summary(HAB_MRIP_small$brevis_count)

#ggplot(HAB_MRIP_small, aes(x=catch_count)) + geom_histogram(binwidth=.5)

remove_outliers <- HAB_MRIP_small %>% filter(catch_count < 600,brevis_count > 0)
head(remove_outliers)


#http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r

#Plot Correlation coeffecient

#http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r

ggscatter(remove_outliers, x = "mean_brevis_count", y = "catch_count", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Mean K. brevis", ylab = "Game Fish Caught", title="Pearson's Correllation - Catch vs Brevis",
          add.params = list(color = "blue", fill = "lightgray"),cor.coef.coord =c(4999000,500))
```

```{r}
head(HAB_MRIP_small)
medium_to_high_brevis <- HAB_MRIP_small %>% filter(mean_brevis_count > 100000)
head(medium_to_high_brevis)

#http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r

ggscatter(medium_to_high_brevis, x = "brevis_count", y = "catch_count", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "K. brevis", ylab = "Game Fish Caught")

```

